---
title: "Compas - Data Cleaning"
author: "Simon Schrauth"
date: "`r Sys.Date()`"
output: html_document
---

# Clean Compas Dataset

## Goal

After executing this script, the compas dataset should be cleaned and ready for modelling in Python.

## Install packages

```{r}
pacman::p_load(tidyverse, 
               here,
               naniar,
               reticulate,
               lubridate,
               wesanderson,
               ggdist)
```

## Load data

```{r}
compas_data_raw = readRDS(here("02_compas", "01_compas_data", "02_compas_data_processed",
                               "01_compas_data_recid_load.rds"))
```

## Glimpse on the data

### Overview over data

```{r}
compas_data_raw
```

```{r}
str(compas_data_raw)
```

### Check for Missing Values

Table of absolute missing values per variable as well as percentage of missing values per variable

```{r}
miss_var_summary(compas_data_raw)
```

Plot with most frequent missingn patterns with according absolute frequencies

```{r}
gg_miss_upset(compas_data_raw) 
```

## Cleaning

- delete patient_nbr and make encounter_id the rowname 
- delete weight, payer_code and medical_specialty, examide, citoglipton, metformin-rosiglitazone, metformin-pioglitazone
- delete incomplete cases (wrt race, diag_1, diag_2, diag_2) = 3713 cases
- delete cases with "no" in readmitted variable (no record of readmission) = 52338 cases
=> 45,715 cases in final dataset

- change readmitted into classes (with "<30" as positive class) -> rate of 1:3.13 (p:n)?
- change data type appropriately (factors with python?)

```{r}
compas_data_clean = compas_data_raw %>% 
  arrange(id) %>% 
  column_to_rownames(var = "id") %>% 
  select(sex, 
         race,
         age, 
         #age_cat,
         juv_fel_count,
         juv_misd_count,
         juv_other_count,
         priors_count = priors_count...15, 
         charge_degree = c_charge_degree, 
         days_b_screening_arrest, 
         is_recid, 
         c_jail_in, 
         c_jail_out,
         #decile_score = decile_score...12,
         two_year_recid) %>% 
  filter(days_b_screening_arrest <= 30,
         days_b_screening_arrest >= -30,
         is_recid != -1,
         charge_degree != "O") %>% 
  mutate(c_jail_in = floor_date(c_jail_in, unit = "day"),
         c_jail_out = floor_date(c_jail_out, unit = "day")) %>% 
  mutate(days_in_jail = as.numeric(round(difftime(c_jail_out, 
                                                  c_jail_in, 
                                                  unit = "days")
                                         )
                                   )
         ) %>% 
  mutate(across(.cols = c(sex,
                          race,
                          charge_degree),
                .fns = ~ factor(.))) %>% 
  select(!c(days_b_screening_arrest,
            c_jail_in,
            c_jail_out,
            is_recid)) %>%
  select(sex,
         race,
         age,
         days_in_jail,
         juv_fel_count,
         juv_misd_count,
         juv_other_count,
         priors_count,
         charge_degree,
         two_year_recid)
```


## Exploratory Data Analysis
### Dataset overview
```{r}
compas_data_clean
```

```{r}
str(compas_data_clean)
```

### Visualization of the class variable (readmitted) & the protected variables (race, age & gender)

#### Definition of colors
```{r}
colors_discrete = rep(wes_palette("Darjeeling1"), 3)
```

#### Class variable: Recidivism
```{r}
compas_eda_class = compas_data_clean %>% 
  select(two_year_recid) %>% 
  mutate(two_year_recid = ifelse(two_year_recid == 1, "yes", "no")) %>% 
  ggplot(aes(x = fct_rev(fct_infreq(two_year_recid)), fill = fct_rev(fct_infreq(two_year_recid)))) +
  geom_bar() +
  labs(title = "Class Frequencies in the Compas Dataset",
       subtitle = "With Class Variable 'Recidivism after two years'",
       x = "Recidivism", 
       y = "Count") +
  scale_fill_manual(values = colors_discrete) +
  guides(fill = "none")

compas_eda_class
```

```{r}
ggsave("01_compas_eda_class.png",
       plot = compas_eda_class,
       device = "png",
       path = here("02_compas", "03_compas_outputs", "03_compas_outputs_graphs"),
       dpi = 300,
       width = 25,
       height = 15,
       units = c("cm")
       )
```

#### Protected variable: Race
```{r}
compas_eda_race = compas_data_clean %>% 
  select(race) %>% 
  ggplot(aes(x = fct_infreq(race), fill = fct_infreq(race))) +
  geom_bar() +
  labs(title = "Race Frequencies in the Compas Dataset",
       x = "Race", 
       y = "Count") +
  scale_fill_manual(values = colors_discrete) +
  guides(fill = "none")

compas_eda_race
```

```{r}
ggsave("02_compas_eda_race.png",
       plot = compas_eda_race,
       device = "png",
       path = here("02_compas", "03_compas_outputs", "03_compas_outputs_graphs"),
       dpi = 300,
       width = 25,
       height = 15,
       units = c("cm")
       )
```

#### Protected variable: Age
```{r}
compas_eda_age_raincloud = compas_data_clean %>% 
  select(age) %>% 
  ggplot(aes(x = age, fill = factor(1))) +
  ggdist::stat_halfeye(
    adjust = 0.75,
    justification = -.2,
    .width = 0,
    point_colour = NA
  ) +
  geom_boxplot(
    width = .12,
    outlier.colour = NA,
    alpha = 0.5
  ) +
  labs(title = "Age Distribution in the Compas Dataset",
       subtitle = "Probability Density Function & Boxplot",
       x = "Age", 
       y = "") +
  scale_fill_manual(values = colors_discrete[3]) +
  guides(fill = "none") +
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank())

compas_eda_age_raincloud
```

```{r}
compas_eda_age_hist = compas_data_clean %>% 
  select(age) %>% 
  ggplot(aes(x = age, fill = factor(1))) +
  geom_histogram(binwidth = 2) +
  labs(title = "Age Distribution in the Compas Dataset",
       subtitle = "Histogram with Binwidth 2",
       x = "Age", 
       y = "Density") +
  scale_fill_manual(values = colors_discrete[3]) +
  guides(fill = "none") 

compas_eda_age_hist
```

```{r}
ggsave("03_compas_eda_age_raincloud.png",
       plot = compas_eda_age_raincloud,
       device = "png",
       path = here("02_compas", "03_compas_outputs", "03_compas_outputs_graphs"),
       dpi = 300,
       width = 25,
       height = 15,
       units = c("cm")
       )

ggsave("04_compas_eda_age_hist.png",
       plot = compas_eda_age_hist,
       device = "png",
       path = here("02_compas", "03_compas_outputs", "03_compas_outputs_graphs"),
       dpi = 300,
       width = 25,
       height = 15,
       units = c("cm")
       )
```

#### Protected variable: Gender
```{r}
compas_eda_sex = compas_data_clean %>% 
  select(sex) %>% 
  ggplot(aes(x = fct_infreq(sex), fill = fct_infreq(sex))) +
  geom_bar() +
  labs(title = "Sex Frequencies in the Compas Dataset",
       x = "Sex", 
       y = "Count") +
  scale_fill_manual(values = colors_discrete) +
  guides(fill = "none")

compas_eda_sex
```

```{r}
ggsave("05_compas_eda_sex.png",
       plot = compas_eda_sex,
       device = "png",
       path = here("02_compas", "03_compas_outputs", "03_compas_outputs_graphs"),
       dpi = 300,
       width = 15,
       height = 15,
       units = c("cm")
       )
```


#### Protected variables: Race & Gender
```{r}
compas_eda_race_sex = compas_data_clean %>% 
  select(race, sex) %>% 
  ggplot(aes(x = sex, fill = sex)) +
  geom_bar() +
  facet_wrap(~race) +
  labs(title = "Sex Frequencies in the Compas Dataset",
       subtitle = "Grouped by Race",
       x = "Sex", 
       y = "Count") +
  scale_fill_manual(values = colors_discrete) +
  guides(fill = "none") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

compas_eda_race_sex
```

```{r}
ggsave("06_compas_eda_race_sex.png",
       plot = compas_eda_race_sex,
       device = "png",
       path = here("02_compas", "03_compas_outputs", "03_compas_outputs_graphs"),
       dpi = 300,
       width = 25,
       height = 15,
       units = c("cm")
       )
```



#### Protected variables: Race & Age
```{r}
compas_eda_race_age = compas_data_clean %>% 
  select(age, race) %>% 
  ggplot(aes(x = race, y = age, fill = race)) +
  ggdist::stat_halfeye(
    adjust = 0.75,
    justification = -.2,
    .width = 0,
    point_colour = NA
  ) +
  geom_boxplot(
    width = .12,
    outlier.colour = NA,
    alpha = 0.5
  ) +
  labs(title = "Age Distribution in the Compas Dataset",
       subtitle = "Grouped by Race",
       x = "Race", 
       y = "Age") +
  scale_fill_manual(values = colors_discrete) +
  guides(fill = "none") +
  coord_flip()

compas_eda_race_age
```

```{r}
ggsave("07_compas_eda_race_age.png",
       plot = compas_eda_race_age,
       device = "png",
       path = here("02_compas", "03_compas_outputs", "03_compas_outputs_graphs"),
       dpi = 300,
       width = 25,
       height = 15,
       units = c("cm")
       )
```


#### Protected variables: Age & Gender
```{r}
compas_eda_sex_age = compas_data_clean %>% 
  select(age, sex) %>% 
  ggplot(aes(x = sex, y = age, fill = sex)) +
  ggdist::stat_halfeye(
    adjust = 0.75,
    justification = -.09,
    .width = 0,
    point_colour = NA
  ) +
  geom_boxplot(
    width = .12,
    outlier.colour = NA,
    alpha = 0.5
  ) +
  labs(title = "Age Distribution in the Compas Dataset",
       subtitle = "Grouped by Sex",
       x = "Sex", 
       y = "Age") +
  scale_fill_manual(values = colors_discrete) +
  guides(fill = "none") +
  coord_flip()

compas_eda_sex_age
```

```{r}
ggsave("08_compas_eda_sex_age.png",
       plot = compas_eda_sex_age,
       device = "png",
       path = here("02_compas", "03_compas_outputs", "03_compas_outputs_graphs"),
       dpi = 300,
       width = 25,
       height = 15,
       units = c("cm")
       )
```


## Save data

```{r}
saveRDS(compas_data_clean, 
        file = here("02_compas", "01_compas_data", "02_compas_data_processed",
                    "03_compas_data_clean.rds")
        )
```



