---
title: "Creditcard - Data Cleaning"
author: "Simon Schrauth"
date: "`r Sys.Date()`"
output: html_document
---

# Clean Creditcard Dataset

## Goal

After executing this script, the creditcard dataset should be cleaned and ready for modelling in Python.

## Install packages

```{r}
pacman::p_load(tidyverse, 
               here,
               naniar,
               reticulate,
               wesanderson,
               ggdist,
               tidyquant)
```

## Load data

```{r}
cc_data_raw = readRDS(file = here("03_creditcard", "01_creditcard_data", "02_creditcard_data_processed",
                                  "01_creditcard_data_load.rds"))
```

## Glimpse on the data

### Overview over data

```{r}
cc_data_raw
```

```{r}
str(cc_data_raw)
```

### Check for Missing Values

Table of absolute missing values per variable as well as percentage of missing values per variable

```{r}
miss_var_summary(cc_data_raw)
```

## Cleaning

- make ID the rowname 
- no missing values => 30000 cases in final dataset
- relevel sex, education & marriage variables

- change data type appropriately (factors with python?)

```{r}
cc_data_clean = cc_data_raw %>% 
  arrange(ID) %>% 
  column_to_rownames(var = "ID") %>% 
  rename(default_payment = "default payment next month") %>% 
  rename_with(tolower) %>% 
  mutate(sex = fct_recode(as.character(sex), 
                          male = "1",
                          female = "2"),
         education = fct_recode(as.character(education),
                                 graduateschool = "1",
                                 university = "2",
                                 highschool = "3",
                                 others = "0",
                                 others = "4",
                                 others = "5",
                                 others = "6"),
         marriage = fct_recode(as.character(marriage),
                               married = "1",
                               single = "2",
                               others = "0",
                               others = "3")        
        )
```

## Exploratory Data Analysis
### Dataset overview
```{r}
cc_data_clean
```

```{r}
str(cc_data_clean)
```

### Visualization of the class variable (readmitted) & the protected variables (race, age & gender)

#### Definition of colors
```{r}
colors_discrete = rep(wes_palette("Darjeeling1"), 3)
```

#### Class variable: Default Payment
```{r}
cc_eda_class = cc_data_clean %>% 
  select(default_payment) %>% 
  mutate(default_payment = ifelse(default_payment == 1, "yes", "no")) %>% 
  ggplot(aes(x = fct_rev(fct_infreq(default_payment)), fill = fct_rev(fct_infreq(default_payment)))) +
  geom_bar() +
  labs(title = "Class Frequencies in the Creditcard Dataset",
       subtitle = "With Class Variable 'Default Payment in the next month'",
       x = "Default Payment in the next month", 
       y = "Count") +
  scale_fill_manual(values = colors_discrete) +
  guides(fill = "none")

cc_eda_class
```

```{r}
ggsave("01_cc_eda_class.png",
       plot = cc_eda_class,
       device = "png",
       path = here("03_creditcard", "03_creditcard_outputs", "03_creditcard_outputs_graphs"),
       dpi = 300,
       width = 25,
       height = 15,
       units = c("cm")
       )
```

#### Protected variable: Sex
```{r}
cc_eda_sex = cc_data_clean %>% 
  select(sex) %>% 
  ggplot(aes(x = fct_infreq(sex), fill = fct_infreq(sex))) +
  geom_bar() +
  labs(title = "Sex Frequencies in the Creditcard Dataset",
       x = "Sex", 
       y = "Count") +
  scale_fill_manual(values = colors_discrete) +
  guides(fill = "none")

cc_eda_sex
```

```{r}
ggsave("02_cc_eda_sex.png",
       plot = cc_eda_sex,
       device = "png",
       path = here("03_creditcard", "03_creditcard_outputs", "03_creditcard_outputs_graphs"),
       dpi = 300,
       width = 25,
       height = 15,
       units = c("cm")
       )
```

#### Protected Variable: Marriage
```{r}
cc_eda_marriage = cc_data_clean %>% 
  select(marriage) %>% 
  ggplot(aes(x = fct_infreq(marriage), fill = fct_infreq(marriage))) +
  geom_bar() +
  labs(title = "Frequencies of Marriage Status in the Creditcard Dataset",
       x = "Marriage Status", 
       y = "Count") +
  scale_fill_manual(values = colors_discrete) +
  guides(fill = "none")

cc_eda_marriage
```
```{r}
ggsave("03_cc_eda_marriage.png",
       plot = cc_eda_marriage,
       device = "png",
       path = here("03_creditcard", "03_creditcard_outputs", "03_creditcard_outputs_graphs"),
       dpi = 300,
       width = 25,
       height = 15,
       units = c("cm")
       )
```

#### Protected variable: Age
```{r}
cc_eda_age_raincloud = cc_data_clean %>% 
  select(age) %>% 
  ggplot(aes(x = age, fill = factor(1))) +
  ggdist::stat_halfeye(
    adjust = 0.75,
    justification = -.2,
    .width = 0,
    point_colour = NA
  ) +
  geom_boxplot(
    width = .12,
    outlier.colour = NA,
    alpha = 0.5
  ) +
  labs(title = "Age Distribution in the Creditcard Dataset",
       subtitle = "Probability Density Function & Boxplot",
       x = "Age", 
       y = "") +
  scale_fill_manual(values = colors_discrete[3]) +
  guides(fill = "none") +
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank())

cc_eda_age_raincloud
```

```{r}
cc_eda_age_hist = cc_data_clean %>% 
  select(age) %>% 
  ggplot(aes(x = age, fill = factor(1))) +
  geom_histogram(binwidth = 2) +
  labs(title = "Age Distribution in the Creditcard Dataset",
       subtitle = "Histogram with Binwidth 2",
       x = "Age", 
       y = "Density") +
  scale_fill_manual(values = colors_discrete[3]) +
  guides(fill = "none") 

cc_eda_age_hist
```


```{r}
ggsave("04_cc_eda_age_raincloud.png",
       plot = cc_eda_age_raincloud,
       device = "png",
       path = here("03_creditcard", "03_creditcard_outputs", "03_creditcard_outputs_graphs"),
       dpi = 300,
       width = 25,
       height = 15,
       units = c("cm")
       )

ggsave("05_cc_eda_age_hist.png",
       plot = cc_eda_age_hist,
       device = "png",
       path = here("03_creditcard", "03_creditcard_outputs", "03_creditcard_outputs_graphs"),
       dpi = 300,
       width = 25,
       height = 15,
       units = c("cm")
       )
```

#### Protected variables: Sex & Marriage Status
```{r}
cc_eda_sex_marriage = cc_data_clean %>% 
  select(sex, marriage) %>% 
  ggplot(aes(x =  fct_infreq(marriage), fill = fct_infreq(marriage))) +
  geom_bar() +
  facet_wrap(~sex) +
  labs(title = "Frequencies of Marriage Status in the Creditdata Dataset",
       subtitle = "Grouped by Sex",
       x = "Marriage Status", 
       y = "Count") +
  scale_fill_manual(values = colors_discrete) +
  guides(fill = "none") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

cc_eda_sex_marriage
```

```{r}
ggsave("06_cc_eda_sex_marriage.png",
       plot = cc_eda_sex_marriage,
       device = "png",
       path = here("03_creditcard", "03_creditcard_outputs", "03_creditcard_outputs_graphs"),
       dpi = 300,
       width = 25,
       height = 15,
       units = c("cm")
       )
```

#### Protected variables: Sex & Age
```{r}
cc_eda_marriage_age = cc_data_clean %>% 
  select(age, marriage) %>% 
  ggplot(aes(x = marriage, y = age, fill = marriage)) +
  ggdist::stat_halfeye(
    adjust = 0.75,
    justification = -.2,
    .width = 0,
    point_colour = NA
  ) +
  geom_boxplot(
    width = .12,
    outlier.colour = NA,
    alpha = 0.5
  ) +
  labs(title = "Age Distribution in the Creditcard Dataset",
       subtitle = "Grouped by Marriage Status",
       x = "Marriage Status", 
       y = "Age") +
  scale_fill_manual(values = colors_discrete) +
  guides(fill = "none") +
  coord_flip()

cc_eda_marriage_age
```

```{r}
ggsave("07_cc_eda_marriage_age.png",
       plot = cc_eda_marriage_age,
       device = "png",
       path = here("03_creditcard", "03_creditcard_outputs", "03_creditcard_outputs_graphs"),
       dpi = 300,
       width = 25,
       height = 15,
       units = c("cm")
       )

```

## Save data

```{r}
saveRDS(cc_data_clean, 
        file = here("03_creditcard", "01_creditcard_data", "02_creditcard_data_processed",
                    "02_creditcard_data_clean.rds")
        )
```



